---
title: "Expoloring the data from the CliRes database"
output:
  html_document:
    theme: cosmo
    toc: true
editor_options: 
  chunk_output_type: console
css: style.css
---

```{r general_options, include = FALSE}
knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  prompt = function(before, options, envir) {
    options(prompt = if (options$engine %in% c("sh", "bash")) "$ " else "> ")
  })

knitr::opts_chunk$set(margin = TRUE, prompt = TRUE, comment = "",
                      collapse = TRUE, cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
                      fig.width = 4.24725, fig.retina = 2, fig.align = "center")

options(width = 250)
```

## Packages

Installing the required packages:

```{r}
required <- c("dplyr", "magrittr", "readr")
to_install <- setdiff(required, row.names(installed.packages()))
if (length(to_install)) install.packages(to_install)
```

Loading `magrittr`:

```{r}
library(magrittr)
```

## Loading the data

```{r}
viparc <- readr::read_csv("https://raw.githubusercontent.com/viparc/clires_data/master/data/viparc.csv",
                          col_types = paste(c("cii", rep("l", 6), rep("d", 45), "lil"), collapse = ""))
```

## Numbers of farms, flocks and weeks

There are **`r length(unique(viparc$farm))` farms** in the study:

```{r}
length(unique(viparc$farm))
```

And a total of **`r nrow(unique(dplyr::select(viparc, farm, flock)))` flocks**:

```{r}
viparc %>% 
  dplyr::select(farm, flock) %>% 
  unique() %>% 
  nrow()
```

Of which there are **`r nrow(unique(dplyr::select(dplyr::filter(viparc, completed), farm, flock)))` completed flocks** :

```{r}
viparc %>% 
  dplyr::filter(completed) %>% 
  dplyr::select(farm, flock) %>% 
  unique() %>% 
  nrow()
```

This represents **`r nrow(viparc)` weeks** of observation:

```{r}
nrow(viparc)
```

And **`r nrow(dplyr::filter(viparc, completed))` weeks** for completed flocks:

```{r}
nrow(dplyr::filter(viparc, completed))
```

The distribution of the number of flocks per farm:

```{r}
viparc %>% 
  dplyr::select(farm, flock) %>% 
  unique() %>% 
  dplyr::group_by(farm) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$% 
  hist(n, 0:12, col = "grey", main = NA, xlab = "number of flocks", ylab = "number of farms")
```

or:

```{r}
viparc %>% 
  dplyr::select(farm, flock) %>% 
  unique() %>% 
  dplyr::group_by(farm) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$%
  table(n)
```

The same thing, considering only completed flocks:

```{r}
viparc %>% 
  dplyr::filter(completed) %>% 
  dplyr::select(farm, flock) %>% 
  unique() %>% 
  dplyr::group_by(farm) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$% 
  hist(n, 0:12, col = "grey", main = NA, xlab = "number of completed flocks", ylab = "number of farms")
```

or:

```{r}
viparc %>%
  dplyr::filter(completed) %>% 
  dplyr::select(farm, flock) %>% 
  unique() %>% 
  dplyr::group_by(farm) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$%
  table(n)
```

The distribution of the number of weeks per flock:

```{r}
viparc %>% 
  dplyr::group_by(farm, flock) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$% 
  hist(n, 0:27, col = "grey", main = NA, xlab = "number of weeks", ylab = "number of flocks")
```

```{r}
viparc %>% 
  dplyr::group_by(farm, flock) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$% 
  table(n)
```

The same thing, considering only the completed flocks:

```{r}
viparc %>% 
  dplyr::filter(completed) %>% 
  dplyr::group_by(farm, flock) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$% 
  hist(n, 0:27, col = "grey", main = NA, xlab = "number of weeks", ylab = "number of flocks")
```

```{r}
viparc %>% 
  dplyr::filter(completed) %>% 
  dplyr::group_by(farm, flock) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$% 
  table(n)
```

## Number of chicken

The distribution of the flocks sizes:

```{r}
viparc %>% 
  dplyr::filter(week < 2) %$%
  hist(nb_chicken, nclass = 30, col = "grey", main = NA, xlab = "number of chicken", ylab = "number of flocks")
```

Or:

```{r}
viparc %>% 
  dplyr::filter(week < 2) %$%
  head(sort(nb_chicken), 15)
```

The distribution of the farms sizes:

```{r}
viparc %>% 
  dplyr::filter(week < 2) %>%
  dplyr::group_by(farm) %>% 
  dplyr::summarise(size = mean(nb_chicken)) %>%
  dplyr::ungroup() %$% 
  hist(size, col = "grey", main = NA, xlab = "number of chicken", ylab = "number of farms")
```

Or:

```{r}
viparc %>% 
  dplyr::filter(week < 2) %>%
  dplyr::group_by(farm) %>% 
  dplyr::summarise(size = mean(nb_chicken)) %>%
  dplyr::ungroup() %$% 
  head(round(sort(size)), 15)
```

## Feces samplings

Not all the flocks are sampled 3 times:

```{r}
(samplings <- viparc %>% 
  dplyr::select(farm, flock, completed, sampling) %>% 
  dplyr::group_by(farm, flock) %>%
  dplyr::summarise(completed = mean(completed), sampling = sum(sampling, na.rm = TRUE)) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(completed = completed > 0))
```

```{r}
with(samplings, table(completed, sampling))
```

The reasons for less than 3 samplings seems to be

* the flock is uncomplete
* the flock died prematuraly
* some samplings were missed

Let's explore a bit more by focusing on the completed flocks and see whether
the flocks with less than 3 samplings tend to be short ones (i.e. suggesting
premature death of the flock):

```{r}
hist2 <- function(x, ...) hist(x, breaks = 0:30, ...)

# Let's plot the durations of the completed flocks:

viparc %>%
  dplyr::filter(completed) %>% 
  dplyr::group_by(farm, flock) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %$%
  hist2(n, col = "grey", xlab = "duration (weeks)", ylab = "number of flocks", main = NA)

# Let's now plot the durations of the completed flocks with 1 sample only:

samplings %>% 
  dplyr::filter(completed, sampling == 1) %$% 
  purrr::map2(farm, flock, function(x, y) nrow(dplyr::filter(viparc, farm == x, flock == y))) %>% 
  unlist() %>% 
  hist2(col = adjustcolor("red", .5), add = TRUE)

# And the durations of the completed flocks with 2 samples only:

samplings %>% 
  dplyr::filter(completed, sampling == 2) %$% 
  purrr::map2(farm, flock, function(x, y) nrow(dplyr::filter(viparc, farm == x, flock == y))) %>% 
  unlist() %>% 
  hist2(col = adjustcolor("blue", .5), add = TRUE)

legend("left", c("1 sampling only", "2 samplings only"), fill = adjustcolor(c("red", "blue"), .5), bty = "n")
```
