---
title: "Generating the data from the CliRes database"
output:
  html_document:
    theme: cosmo
    toc: true
editor_options: 
  chunk_output_type: console
css: style.css
---

```{r general_options, include = FALSE}
knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  prompt = function(before, options, envir) {
    options(prompt = if (options$engine %in% c("sh", "bash")) "$ " else "> ")
  })

knitr::opts_chunk$set(margin = TRUE, prompt = TRUE, comment = "",
                      collapse = TRUE, cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
                      fig.width = 4.24725, fig.retina = 2, fig.align = "center")

options(width = 250)
```

## CliRes excel files

The CliRes data are in 2 excel files, each with many tabs. One of these files
contains general information and the other one contains the surveillance data:

```{r}
general_data <- "https://www.dropbox.com/s/3hfnhd39fxjmcnh/20-5-2019-_18ZN_Category_V1_Data.xls?dl=0"
surveillance_data <- "https://www.dropbox.com/s/4kk6qewmntw3q5o/20-5-2019-_18ZN_V1_Data.xls?dl=0"
```

## Packages

Installing the required packages:

```{r}
required <- c("dplyr", "magrittr", "purrr", "readxl", "tibble", "tidyr")
to_install <- setdiff(required, row.names(installed.packages()))
if (length(to_install)) install.packages(to_install)
```

Loading `magrittr`:

```{r}
library(magrittr)
```

## Utilitary functions

Naming output of `lapply()`:

```{r}
lapply2 <- function(X, FUN, ...) setNames(lapply(X, FUN, ...), X)
```

Tuning the `download.file()` function in order to be able to download from a
Dropbox link:

```{r}
download_dropbox <- function(url, ...) download.file(sub("dl=0", "raw=1", url), ...)
```

The following function returns the flocks with missing weeks:

```{r}
missing_weeks <- function(x) {
  tmp <- x %>%
    dplyr::group_by(USUBJID, FLOCKSEQUENCE) %>%
    dplyr::arrange(WEEK) %>%
    dplyr::summarise(a = length(unique(diff(WEEK)))) %>%
    dplyr::ungroup() %>%
    dplyr::filter(a > 1)
  if (nrow(tmp)) return(purrr::map2(tmp$USUBJID,
                                    tmp$FLOCKSEQUENCE,
                                    function(farm, flock)
                                      x %>%
                                        dplyr::filter(USUBJID == farm, FLOCKSEQUENCE == flock) %>% 
                                        dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)))
  message("There is no missing week.")
  invisible(0)
}
```

## ESVAC's IU-to-mg conversion data

The
[ESVAC](https://www.ema.europa.eu/en/veterinary-regulatory/overview/antimicrobial-resistance/european-surveillance-veterinary-antimicrobial-consumption-esvac)
conversion factors are available from the table 2 of page 62 of this
[OIE](http://www.oie.int)
[report](http://www.oie.int/fileadmin/Home/fr/Our_scientific_expertise/docs/pdf/AMR/Survey_on_monitoring_antimicrobial_agents_Dec2016.pdf).
A copy-paste of this table without headers is available
[here](https://raw.githubusercontent.com/viparc/clires_data/master/raw_data/esvac.txt).
Let's download this file:

```{r}
tmp <- tempfile()
download.file("https://raw.githubusercontent.com/viparc/clires_data/master/raw_data/esvac.txt", tmp)
```

And read and reformat the data into a named numeric vector:

```{r}
(esvac <- tmp %>% 
  readLines() %>% 
  sub("^#.*", "", .) %>% 
  sub(" *", "", .) %>% 
  {.[. != ""]} %>% 
  matrix(ncol = 4, byrow = TRUE) %>% 
  as.data.frame(stringsAsFactors = FALSE) %>% 
  setNames(c("ab_vet_med", "ab_oie", "IU.mg", "mg.IU")) %>% 
  dplyr::filter(!grepl("methane", ab_vet_med)) %>% 
  dplyr::transmute(ab_oie = sub(" .*$", "", tolower(ab_oie)),
                   mg.IU  = as.numeric(mg.IU)) %>% 
  dplyr::bind_rows(data.frame(ab_oie = "josamycin",
                              mg.IU  = 0.001, stringsAsFactors = FALSE)) %$%
  setNames(mg.IU, ab_oie))
```

Note that we got rid off one of the colistin data and that we manually added the
data for josamycin that was absent from the table.

## Drug codes from general information data

Downloading the general data file:

```{r}
tmp <- tempfile()
download_dropbox(general_data, tmp)
```

Reformating and merging with the `esvac` dataframe that we generated above:

```{r}
(drug_codes <- c("ANTIMICROBIAL", "ANTIMICROBIAL_GridAnti") %>% 
  lapply(readxl::read_excel, path = tmp) %>% 
  c(list(c("ANTIMICROBIAL_SEQ"))) %>% 
  do.call(dplyr::right_join, .) %>%
  dplyr::select(CODE, ANTINAME1, CONCENTRATION, GRIDUNIT, GRIDPACK, GRIDPACKUNIT) %>%
  dplyr::filter(! ANTINAME1 %in% c("alicin", "axit oxolinic", "iodo-hydroxyquinoline", "metronidazol", "nystatin")) %>% 
  dplyr::mutate(ANTINAME1     = dplyr::recode(ANTINAME1, sunfadimethoxine  = "sulfadimethoxine",
                                                         sunphamethoxazole = "sulphamethoxazole"),
                GRIDUNIT      = dplyr::na_if(GRIDUNIT, "NA"),
                GRIDUNIT      = dplyr::na_if(GRIDUNIT, "na"),
                GRIDUNIT      = dplyr::recode(GRIDUNIT, UI = "IU", mg = "MG"),
                GRIDPACKUNIT  = dplyr::na_if(GRIDPACKUNIT, "na"),
                GRIDPACKUNIT  = dplyr::na_if(GRIDPACKUNIT, "VI"),
                GRIDPACKUNIT  = dplyr::recode(GRIDPACKUNIT, g  = "G", ml = "G", ML = "G"),
                GRIDPACK      = dplyr::na_if(GRIDPACK, "na"),
                GRIDPACK      = as.numeric(GRIDPACK),
                CONCENTRATION = dplyr::na_if(CONCENTRATION, "NA"),
                CONCENTRATION = dplyr::na_if(CONCENTRATION, "na"),
                CONCENTRATION = dplyr::recode(CONCENTRATION, S = "500"), # this will ultimately be corrected in CliRes CORRECTION
                CONCENTRATION = as.numeric(CONCENTRATION),
                CONCENTRATION = ifelse(GRIDUNIT == "IU", CONCENTRATION * esvac[ANTINAME1], CONCENTRATION),
                GRIDUNIT      = ifelse(GRIDUNIT == "IU", "MG", GRIDUNIT), # linked to the above line
                CONCENTRATION = ifelse(GRIDUNIT == "MG", CONCENTRATION / 1000, CONCENTRATION),
                GRIDUNIT      = ifelse(GRIDUNIT == "MG", "G", GRIDUNIT), # linked to the above line
                GRIDPACK      = ifelse(GRIDPACKUNIT == "KG", 1000 * GRIDPACK, GRIDPACK),
                GRIDPACKUNIT  = ifelse(GRIDPACKUNIT == "KG", "G", GRIDPACKUNIT), # linked to the above line
                proportion    = CONCENTRATION / GRIDPACK) %>% 
  dplyr::select(CODE, ANTINAME1, proportion))
```

## The surveillance data

Downloading the surveillance data file:

```{r}
tmp <- tempfile()
download_dropbox(surveillance_data, tmp)
```

For antimicrobial, disease and chicken data in the surveillance file, we need
the following pairs of tabs respectively, that we have to merge:

```{r}
antimicrobial <- c("ANTIMICROBIAL", "ANTIMICROBIAL_GridEvent")
disease <- c("DISEASE", "DISEASE_GridEvent")
chicks <- c("MID_INOUT", "MID_INOUT_GridEvent")
```

Reading the data from file:

```{r}
(surveillance <- lapply2(c(antimicrobial, disease, chicks, "SAMPLE"),
                         readxl::read_excel, path = tmp))
```

### Sampling dates

Retrieving the sampling dates. One key feature of the code below is to convert
a sampling date into a week index. `SAMPLINGVISIT` tells whether the visit is at
the start (`S`), the middle (`M`) or the end (`E`) of the flock.

```{r}
(samples <- surveillance$SAMPLE %>%
# End sampling of flock 2 of farm 75-013 is duplicated, so we filter it out: CORRECTION
  dplyr::filter(!(USUBJID == "75-013" & FLOCKSEQUENCE == "02" & SAMPLE_SEQ == 8)) %>% 
  dplyr::select(USUBJID, FLOCKSEQUENCE, SAMPLINGDATE, SAMPLINGVISIT) %>%
  assign("tmp", ., 1) %>% # for reuse 4 lines later
  dplyr::filter(SAMPLINGVISIT == "S") %>% # | gets the date...
  dplyr::select(-SAMPLINGVISIT) %>%       # | ... of the first week...
  dplyr::rename(start = SAMPLINGDATE) %>% # | ... of each flock.
  dplyr::right_join(tmp, c("USUBJID", "FLOCKSEQUENCE")) %>%
  dplyr::mutate(WEEK     = as.numeric(floor((SAMPLINGDATE - start) / (60 * 60 * 24 * 7)) + 1),
                sampling = TRUE) %>% 
  dplyr::select(-start, -SAMPLINGDATE, -SAMPLINGVISIT) %>% 
  unique()) # dates of first visits are missing for flock 2 of farm 75-019, THIS NEED TO BE FIXED!!!!!
            # flock 3 of farm 75-029 and flock 2 of farm 75-058. CORRECTION
```

This data frame contains all the dates of sampling, one per row.

### Clinical signs

Retrieving the clinical signs data. Note that there are 2 corrections that we
need to do here.

```{r}
(clinical_signs <- surveillance[disease] %>% 
  unname() %>% # because of the do.call() call that follows
  c(list(c("USUBJID", "DISEASE_SEQ"))) %>% 
  do.call(dplyr::left_join, .) %>%
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK, RESPIRATORY, DIARRHOEA, CNS, MALAISE, LEGLESIONS, SUDDENDEATH, DISEASE_GridEvent_SEQ) %>% 
  assign("tmp", ., 1) %>% # for reuse 3 lines later
# CORRECTION 1: weeks 8 instead of 9 for one of the two records of cycle 2 of farm 8:
  dplyr::filter(USUBJID == "75-008", FLOCKSEQUENCE == "02", DISEASE_GridEvent_SEQ == 9) %>% 
  dplyr::mutate(WEEK = 9) %>%
  dplyr::bind_rows(dplyr::filter(tmp, !(USUBJID == "75-008" & FLOCKSEQUENCE == "02" & DISEASE_GridEvent_SEQ == 9))) %>% 
  assign("tmp", ., 1) %>% # for reuse 3 lines later
# CORRECTION 2: week 19 instead of 10 for cycle 6 of farm 21:
  dplyr::filter(USUBJID == "75-021", FLOCKSEQUENCE == "06", DISEASE_GridEvent_SEQ == 10) %>% 
  dplyr::mutate(WEEK = 10) %>%
  dplyr::bind_rows(dplyr::filter(tmp, !(USUBJID == "75-021" & FLOCKSEQUENCE == "06" & DISEASE_GridEvent_SEQ == 10))) %>% 
  dplyr::select(-DISEASE_GridEvent_SEQ))
```

This data frame contains all the weeks of all the flocks of all the farms. There
is no missing weeks:

```{r}
missing_weeks(clinical_signs)
```

### Antimicrobial usage

Here we'll merge with the `drug_codes` dataframe that we generated above.

```{r}
(amu <- surveillance[antimicrobial] %>% 
  unname() %>% # because of the do.call() call that follows
  c(list(c("USUBJID", "ANTIMICROBIAL_SEQ"))) %>% 
  do.call(dplyr::left_join, .) %>%
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEKNO, CODE, AMOUTUSED, PACKAGEUNIT) %>%
  dplyr::rename(WEEK = WEEKNO) %>% 
  dplyr::mutate(PACKAGEUNIT = dplyr::na_if(PACKAGEUNIT, "TAB")) %>% 
  dplyr::left_join(drug_codes, "CODE") %>%
  tidyr::replace_na(list(ANTINAME1 = "unknown")) %>% # because some antibiotic names are unknown
  dplyr::mutate(antibiotic  = proportion * AMOUTUSED,
                ANTINAME1   = paste0(ANTINAME1, "_g")) %>%
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK, ANTINAME1, antibiotic) %>% 
  tibble::rowid_to_column() %>% # spread() requires unique row identifiers
  tidyr::spread(ANTINAME1, antibiotic) %>%
  dplyr::select(-rowid) %>% # we don't need rowid anymore
  replace(is.na(.), 0) %>%
  dplyr::group_by(USUBJID, FLOCKSEQUENCE, WEEK) %>%
  dplyr::summarise_all(~sum(.)) %>%
  dplyr::ungroup() %>% 
  dplyr::mutate_at(dplyr::vars(dplyr::ends_with("_g")), . %>% ifelse(is.na(.), 0, .)))
```

No missing weeks:

```{r}
missing_weeks(amu)
```

### Chicken data

Below is the code to retrieve data on chicken demographics. In there, we are
computing a variable indicating whether the cycle is finisheed.

```{r}
(chicken <- surveillance[chicks] %>% 
  unname() %>% # because of the do.call() call that follows
  c(list(c("USUBJID", "MID_INOUT_SEQ"))) %>% 
  do.call(dplyr::left_join, .) %>%
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK, CHICKENTOTAL, MID_INOUT_GridEvent_SEQ) %>% 
# CORRECTION:
  assign("tmp", ., 1) %>% # for reuse 3 lines below
  dplyr::filter(USUBJID == "75-070", FLOCKSEQUENCE == "08", MID_INOUT_GridEvent_SEQ == 6) %>% 
  dplyr::mutate(WEEK = 6) %>% 
  dplyr::bind_rows(dplyr::filter(tmp, !(USUBJID == "75-070" & FLOCKSEQUENCE == "08" & MID_INOUT_GridEvent_SEQ == 6))) %>% 
  dplyr::select(-MID_INOUT_GridEvent_SEQ) %>% 
# (end of correction)
  dplyr::mutate_at("CHICKENTOTAL", as.integer) %>% 
  assign("tmp", ., 1) %>% # for reuse on the final line
  dplyr::group_by(USUBJID, FLOCKSEQUENCE) %>%             # | computing whether...
  dplyr::summarize(completed = min(CHICKENTOTAL) < 1) %>% # | ... the cycle is...
  dplyr::ungroup() %>%                                    # | ... finished.
  dplyr::right_join(tmp, c("USUBJID", "FLOCKSEQUENCE")))
```

No missing week:

```{r}
missing_weeks(chicken)
```

## Merging into one single data frame

Here, we merge the 4 data frames that were computed above: `samples`,
`clinical_signs`, `amu` and `chicken`. First, we need to check that there is no
incompatibility between the `amu`, `clinical_signs` and `chicken` data frames:

```{r}
dplyr::anti_join(amu, clinical_signs, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)
dplyr::anti_join(clinical_signs, amu, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)
dplyr::anti_join(chicken, clinical_signs, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)
dplyr::anti_join(clinical_signs, chicken, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)
dplyr::anti_join(amu, chicken, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)
dplyr::anti_join(chicken, amu, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)
```

The week `17` of flock `05` of the farm `75-004` should be removed from the `amu`
data set:

```{r}
amu %>%
  dplyr::filter(USUBJID == "75-004", FLOCKSEQUENCE == "05", WEEK == 17) %>%
  dplyr::select(-USUBJID, -FLOCKSEQUENCE, -WEEK) %>%
  unlist() %>%
  unique()
```

Let's remove it:

```{r}
amu %<>% dplyr::filter(!(USUBJID == "75-004" & FLOCKSEQUENCE == "05" & WEEK == 17))
```

Furthermore, is seems that there is no data for the week `18` of the flock `06`
of the farm `75-077` in the `amu` and `clinical_signs` data frame.

```{r}
nrow(amu)
nrow(chicken)
nrow(clinical_signs)
```

We are ready for the merging and writing to disk:

```{r}
if (!dir.exists("data")) dir.create("data")
list(chicken, clinical_signs, amu, samples) %>% # chicken needs to be first because it's the only df with no NA
  purrr::reduce(dplyr::left_join, by = c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>%
  dplyr::mutate(sampling = ! is.na(sampling)) %>% # because in samples there are only sampled weeks
  dplyr::mutate_at(dplyr::vars(FLOCKSEQUENCE, WEEK), as.integer) %>% 
  dplyr::arrange(USUBJID, FLOCKSEQUENCE, WEEK)  %>%
  dplyr::rename(farm         = USUBJID,
                flock        = FLOCKSEQUENCE,
                week         = WEEK,
                respiratory  = RESPIRATORY,
                diarrhoea    = DIARRHOEA,
                cns          = CNS,
                malaise      = MALAISE,
                leg_lesions  = LEGLESIONS,
                sudden_death = SUDDENDEATH,
                nb_chicken   = CHICKENTOTAL) %>% 
  dplyr::select(farm, flock, week, sampling, dplyr::everything()) %>% 
  write.csv("data/viparc.csv", FALSE, row.names = FALSE)
```

Note that, among the `clinical_signs`, `amu`, `chicken` and `samples` data frames,
`samples` is the only one that contains rows only for the weeks where samples
are taken. The three other data frames contain rows for all the weeks of all the
flocks of all the farms. For this reason, after the merging, we have to replace
the missing values in the `sampling` variable by `FALSE`.