---
title: "Generating the data from the CliRes database"
output:
  html_document:
    theme: cosmo
    toc: true
editor_options: 
  chunk_output_type: console
css: style.css
---

```{r general_options, include = FALSE}
knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  prompt = function(before, options, envir) {
    options(prompt = if (options$engine %in% c("sh", "bash")) "$ " else "> ")
  })

knitr::opts_chunk$set(margin = TRUE, prompt = TRUE, comment = "",
                      collapse = TRUE, cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
                      fig.width = 4.24725, fig.retina = 2, fig.align = "center")

options(width = 250)
```

## CliRes excel files

The CliRes data are in 2 excel files, each with many tabs. One of these files
contains general information and the other one contains the surveillance data:

```{r}
general_data <- "https://www.dropbox.com/s/gt6k1m781537vd2/16-4-2019-_18ZN_Category_V1_Data.xls?dl=0"
surveillance_data <- "https://www.dropbox.com/s/7gjt77eltpm4ygw/16-4-2019-_18ZN_V1_Data.xls?dl=0"
```

## Packages

Installing the required packages:

```{r}
required <- c("dplyr", "magrittr", "purrr", "readxl", "tibble", "tidyr")
to_install <- setdiff(required, row.names(installed.packages()))
if (length(to_install)) install.packages(to_install)
```

Loading `magrittr`:

```{r}
library(magrittr)
```

## Utilitary functions

Naming output of `lapply()`:

```{r}
lapply2 <- function(X, FUN, ...) setNames(lapply(X, FUN, ...), X)
```

Tuning the `download.file()` function in order to be able to download from a
Dropbox link:

```{r}
download_dropbox <- function(url, ...) download.file(sub("dl=0", "raw=1", url), ...)
```

The following function returns the flocks with missing weeks:

```{r}
missing_weeks <- function(x) {
  a <- x %>%
    dplyr::group_by(USUBJID, FLOCKSEQUENCE) %>%
    dplyr::arrange(WEEK) %>%
    dplyr::summarise(a = length(unique(diff(WEEK)))) %>%
    dplyr::ungroup() %>%
    dplyr::filter(a > 1)
  if (nrow(a)) return(purrr::map2(a$USUBJID,
                                  a$FLOCKSEQUENCE,
                                  function(farm, flock)
                                    x %>%
                                      dplyr::filter(USUBJID == farm, FLOCKSEQUENCE == flock) %>% 
                                      dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK)
                                    ))
  message("There is no missing week.")
  invisible(0)
}
```

## ESVAC's IU-to-mg conversion data

The
[ESVAC](https://www.ema.europa.eu/en/veterinary-regulatory/overview/antimicrobial-resistance/european-surveillance-veterinary-antimicrobial-consumption-esvac) conversion factors are
available from the table 2 of page 62 of this [OIE](http://www.oie.int) [report](http://www.oie.int/fileadmin/Home/fr/Our_scientific_expertise/docs/pdf/AMR/Survey_on_monitoring_antimicrobial_agents_Dec2016.pdf). A copy-paste of this table without headers is
available
[here](https://raw.githubusercontent.com/viparc/clires_data/master/raw_data/esvac.txt).
Let's download this file:

```{r}
tmp <- tempfile()
download.file("https://raw.githubusercontent.com/viparc/clires_data/master/raw_data/esvac.txt", tmp)
```

And read and reformat the data into a named list:

```{r}
esvac <- tmp %>% 
  readLines() %>% 
  sub("^#.*", "", .) %>% 
  sub(" *", "", .) %>% 
  {.[. != ""]} %>% 
  matrix(ncol = 4, byrow = TRUE) %>% 
  as.data.frame(stringsAsFactors = FALSE) %>% 
  setNames(c("ab_vet_med", "ab_oie", "IU.mg", "mg.IU")) %>% 
  dplyr::filter(!grepl("methane", ab_vet_med)) %>% 
  dplyr::transmute(ab_oie = sub(" .*$", "", tolower(ab_oie)),
                   mg.IU  = as.numeric(mg.IU)) %>% 
  dplyr::bind_rows(data.frame(ab_oie = "josamycin",
                              mg.IU  = 0.001, stringsAsFactors = FALSE)) %$%
  setNames(mg.IU, ab_oie)
```

Note that we got rid off one of the colistin data and that we manually added the
data for josamycin that was absent from the table.

## Drug codes from general information data

Downloading the general data file:

```{r}
tmp <- tempfile()
download_dropbox(general_data, tmp)
```

Reformating and merging with the `esvac` dataframe that we generated above:

```{r}
drug_codes <- c("ANTIMICROBIAL", "ANTIMICROBIAL_GridAnti") %>% 
  lapply(readxl::read_excel, path = tmp) %>% 
  purrr::invoke(dplyr::right_join, ., "ANTIMICROBIAL_SEQ") %>% 
  dplyr::select(CODE, ANTINAME1, CONCENTRATION, GRIDUNIT, GRIDPACK, GRIDPACKUNIT) %>%
  dplyr::filter(! ANTINAME1 %in% c("alicin", "axit oxolinic", "iodo-hydroxyquinoline", "metronidazol", "nystatin")) %>% 
  dplyr::mutate(ANTINAME1     = dplyr::recode(ANTINAME1, sunfadimethoxine  = "sulfadimethoxine",
                                                         sunphamethoxazole = "sulphamethoxazole"),
                GRIDUNIT      = dplyr::na_if(GRIDUNIT, "NA"),
                GRIDUNIT      = dplyr::na_if(GRIDUNIT, "na"),
                GRIDUNIT      = dplyr::recode(GRIDUNIT, UI = "IU", mg = "MG"),
                GRIDPACKUNIT  = dplyr::na_if(GRIDPACKUNIT, "na"),
                GRIDPACKUNIT  = dplyr::na_if(GRIDPACKUNIT, "VI"),
                GRIDPACKUNIT  = dplyr::recode(GRIDPACKUNIT, g  = "G", ml = "G", ML = "G"),
                GRIDPACK      = dplyr::na_if(GRIDPACK, "na"),
                GRIDPACK      = as.numeric(GRIDPACK),
                CONCENTRATION = dplyr::na_if(CONCENTRATION, "NA"),
                CONCENTRATION = dplyr::na_if(CONCENTRATION, "na"),
                CONCENTRATION = dplyr::recode(CONCENTRATION, S = "500"), # this will ultimately be corrected in CliRes
                CONCENTRATION = as.numeric(CONCENTRATION),
                CONCENTRATION = ifelse(GRIDUNIT == "IU", CONCENTRATION * esvac[ANTINAME1], CONCENTRATION),
                GRIDUNIT      = ifelse(GRIDUNIT == "IU", "MG", GRIDUNIT), # linked to the above line
                CONCENTRATION = ifelse(GRIDUNIT == "MG", CONCENTRATION / 1000, CONCENTRATION),
                GRIDUNIT      = ifelse(GRIDUNIT == "MG", "G", GRIDUNIT), # linked to the above line
                GRIDPACK      = ifelse(GRIDPACKUNIT == "KG", 1000 * GRIDPACK, GRIDPACK),
                GRIDPACKUNIT  = ifelse(GRIDPACKUNIT == "KG", "G", GRIDPACKUNIT), # linked to the above line
                proportion    = CONCENTRATION / GRIDPACK)
```

## The surveillance data

Downloading the surveillance data file:

```{r}
tmp <- tempfile()
download_dropbox(surveillance_data, tmp)
```

For antimicrobial, disease and chicken data in the surveillance file, we need
the following pairs of tabs respectively, that we have to merge:

```{r}
antimicrobial <- c("ANTIMICROBIAL", "ANTIMICROBIAL_GridEvent")
disease <- c("DISEASE", "DISEASE_GridEvent")
chicks <- c("MID_INOUT", "MID_INOUT_GridEvent")
```

Reading the data from file:

```{r}
surveillance <- lapply2(c(antimicrobial, disease, chicks, "SAMPLE"),
                        readxl::read_excel, path = tmp)
```

### Sampling dates

Retrieving the sampling dates:

```{r}
samplesdata <- surveillance$SAMPLE %>% 
  dplyr::select(USUBJID, FLOCKSEQUENCE, SAMPLINGDATE, SAMPLINGVISIT) %>% 
  unique()
```

Here we have to break the pipeline because we need to use the `samplesdata` at
two distincts points in the section that follows:

```{r}
samples <- samplesdata %>%
  dplyr::filter(SAMPLINGVISIT == "S") %>%
  dplyr::select(-SAMPLINGVISIT) %>%
  dplyr::rename(start = SAMPLINGDATE) %>%
  dplyr::right_join(samplesdata, c("USUBJID", "FLOCKSEQUENCE")) %>%
  dplyr::mutate(WEEK     = as.numeric(floor((SAMPLINGDATE - start) / (60 * 60 * 24 * 7)) + 1),
                sampling = TRUE) %>% 
  dplyr::select(-start, -SAMPLINGDATE, -SAMPLINGVISIT)
```

This data frame contains all the dates of sampling, one per row.

### Clinical signs

Retrieving the clinical signs data:

```{r}
clinical_signs0 <- surveillance[disease] %>% 
  unname() %>% 
  purrr::invoke(dplyr::left_join, ., c("USUBJID", "DISEASE_SEQ")) %>%
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK, RESPIRATORY, DIARRHOEA, CNS, MALAISE, LEGLESIONS, SUDDENDEATH)
```

There are two corrections that cannot be done in CliRes and that we have to do
here. We also have to break the pipeline because we uses the `clinical_signs0`
dataframe that we just created at two distinct points in the section that
follows:

```{r}
clinical_signs <- clinical_signs0 %>% 
# correction 1: 2 weeks 8 and no week 9 for cycle 2 of farm 8. These 2 weeks are identical
  dplyr::filter(USUBJID == "75-008", FLOCKSEQUENCE == "02", WEEK == 8) %>% 
  dplyr::mutate(WEEK = 8:9) %>%
  dplyr::bind_rows(dplyr::filter(clinical_signs0, !(USUBJID == "75-008" & FLOCKSEQUENCE == "02" & WEEK == 8))) %>% 
# correction 2: week 19 instead of 10 for cycle 6 of farm 21:
  dplyr::mutate(WEEK = ifelse(USUBJID == "75-021" & FLOCKSEQUENCE == "06" & WEEK == 19, 10, WEEK))
```

This data frame contains all the weeks of all the flocks of all the farms. There
is no missing weeks:

```{r}
missing_weeks(clinical_signs)
```

### Antimicrobial usage

Here we'll merge with the `drug_codes` dataframe that we generated above.

```{r}
amu <- surveillance[antimicrobial] %>% 
  unname() %>% 
  purrr::invoke(dplyr::left_join, ., c("USUBJID", "ANTIMICROBIAL_SEQ")) %>% 
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEKNO, CODE, AMOUTUSED, PACKAGEUNIT) %>%
  dplyr::rename(WEEK = WEEKNO) %>% 
  dplyr::mutate(PACKAGEUNIT = dplyr::na_if(PACKAGEUNIT, "TAB")) %>% 
  dplyr::left_join(drug_codes, "CODE") %>%
  tidyr::replace_na(list(ANTINAME1 = "unknown")) %>% # because some antibiotic names are unknown
  dplyr::mutate(presence    = 1,
                antibiotic  = proportion * AMOUTUSED,
                antibiotic2 = antibiotic,
                antiname2   = paste0(ANTINAME1, "_g"),
                antiname3   = paste0(ANTINAME1, "_g.kg"),
                ANTINAME1   = paste0(ANTINAME1, "_use")) %>%
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK, ANTINAME1, presence, antiname2, antibiotic, antiname3, antibiotic2) %>% 
  tibble::rowid_to_column() %>% # spread() requires unique row identifiers
  tidyr::spread(ANTINAME1, presence) %>%
  tidyr::spread(antiname2, antibiotic) %>%
  tidyr::spread(antiname3, antibiotic2) %>%
  dplyr::select(-rowid) %>% # we don't need rowid anymore
  replace(is.na(.), 0) %>%
  dplyr::group_by(USUBJID, FLOCKSEQUENCE, WEEK) %>%
  dplyr::summarise_all(~sum(.)) %>%
  dplyr::mutate_at(dplyr::vars(dplyr::ends_with("_use", FALSE)), function(x) x > 0) %>% 
  dplyr::mutate_at(dplyr::vars(dplyr::ends_with("_use")), . %>% ifelse(is.na(.), FALSE, .)) %>% 
  dplyr::mutate_at(dplyr::vars(dplyr::ends_with("_g")), . %>% ifelse(is.na(.), 0, .)) %>% 
  dplyr::mutate_at(dplyr::vars(dplyr::ends_with("_g.kg")), . %>% ifelse(is.na(.), 0, .))
```

This data frame is supposed to contain one row per week per flock per farm. It
happens that week 16 of flock 3 of farm 81 is missing:

```{r}
missing_weeks(amu)
```

but that won't be a problem when it's merged with `clinical_signs` below.

### Chicken data

Retrieving data on chicken demographics:

```{r}
chickendata <- surveillance[chicks] %>% 
  unname() %>% 
  purrr::invoke(dplyr::left_join, ., c("USUBJID", "MID_INOUT_SEQ")) %>% 
  dplyr::select(USUBJID, FLOCKSEQUENCE, WEEK, CHICKENTOTAL) %>% 
  unique() %>% 
  dplyr::mutate_at("CHICKENTOTAL", as.integer)
```

Here we have to break the pipeline because `chickendata` is used at two distinct
points in the part below:

```{r}
chicken <- chickendata %>%
  dplyr::group_by(USUBJID, FLOCKSEQUENCE) %>%
  dplyr::summarize(completed = min(CHICKENTOTAL) < 1) %>%
  dplyr::ungroup() %>% 
  dplyr::right_join(chickendata, c("USUBJID", "FLOCKSEQUENCE"))
```

This data frame is supposed to contain one row per week per flock per farm. It
happens that week 6 of flock 8 of farm 70 is missing:

```{r}
missing_weeks(chicken)
```

## Merging into one single data frame

Checking that there is no week defined in `amu` but not in `clinical_signs`:

```{r}
dplyr::anti_join(amu, clinical_signs, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% 
  nrow() %>% 
  `<`(1)
```

Checking that there is no week defined in `chicken` but not in `clinical_signs`:

```{r}
dplyr::anti_join(chicken, clinical_signs, c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>% 
  nrow() %>% 
  `<`(1)
```

We are ready for the merging and writing to disk:

```{r}
if (!dir.exists("data")) dir.create("data")
list(clinical_signs, amu, chicken, samples) %>%
  purrr::reduce(dplyr::left_join, by = c("USUBJID", "FLOCKSEQUENCE", "WEEK")) %>%
  dplyr::mutate(sampling = ! is.na(sampling)) %>% 
  dplyr::mutate_at(dplyr::vars(FLOCKSEQUENCE, WEEK), as.integer) %>% 
  dplyr::arrange(USUBJID, FLOCKSEQUENCE, WEEK)  %>%
  dplyr::rename(farm         = USUBJID,
                flock        = FLOCKSEQUENCE,
                week         = WEEK,
                respiratory  = RESPIRATORY,
                diarrhoea    = DIARRHOEA,
                cns          = CNS,
                malaise      = MALAISE,
                leg_lesions  = LEGLESIONS,
                sudden_death = SUDDENDEATH,
                nb_chicken   = CHICKENTOTAL) %>% 
  write.csv("data/viparc.csv", FALSE, row.names = FALSE)
```

Note that, among the `clinical_signs`, `amu`, `chicken` and `samples` data frames,
`samples` is the only one that contains rows only for the weeks where samples
are taken. The three other data frames contain rows for all the weeks of all the
flocks of all the farms. For this reason, after the merging, we have to replace
the missing values in the `sampling` variable by `FALSE`.