---
output: github_document
editor_options: 
  chunk_output_type: console
---

The `make_data.R` R script that generates the data can be generated by the
following command:

```{r eval = FALSE}
# install.packages("knitr")
knitr::purl("https://raw.githubusercontent.com/viparc/clires_data/master/make_data.Rmd", documentation = 1L)
```

The data set, as a flat rectangular CSV file, can be downloaded from
[here](https://raw.githubusercontent.com/viparc/clires_data/master/data/viparc.csv):

```{r eval = FALSE}
# install.packages("readr")
viparc <- readr::read_csv("https://raw.githubusercontent.com/viparc/clires_data/master/data/viparc.csv")
```

The data frame contains 5391 weeks of observation (rows) and 147 variables
(columns). The variables are

* **USUBJID:** farm ID;
* **FLOCKSEQUENCE:** flock ID (in a given farm);
* **WEEK:** week number (in a given flock of a given farm);
* **RESPIRATORY, ..., SUDDENDEATH:** presence/absence of 6 clinical signs in the flock;
* **amoxicillin_use, ..., unknown_use:** presence/absence of 44 antimicrobials
in the flock;
* **amoxicillin_g, ..., unknown_g:** mass, in g, of the 44 antimicriobial used
in the flock;
* **amoxicillin_g.kg, ..., unknown_g.kg:** mass, in g per kg of chicken, of the 44 antimicriobial used in the flock;
* **completed:** boolean informing whether the flock is done (`TRUE`) or still ongoing (`FALSE`);
* **CHICKENTOTAL:** total number of chicken in the farm;
* **sampling:** boolean informing whether there is feces sampling during the
week.