---
output: github_document
editor_options: 
  chunk_output_type: console
---

The `make_data.R` R script that generates the data can be generated by the
following command:

```{r eval = FALSE}
# install.packages("knitr")
knitr::purl("https://raw.githubusercontent.com/viparc/clires_data/master/make_data.Rmd", documentation = 1L)
```

The data set, as a flat rectangular CSV file, can be downloaded from
[here](https://raw.githubusercontent.com/viparc/clires_data/master/data/viparc.csv):

```{r}
# install.packages("readr")
viparc <- readr::read_csv("https://raw.githubusercontent.com/viparc/clires_data/master/data/viparc.csv",
                          col_types = paste(c("cii", rep("l", 6), rep("d", 45), "lil"), collapse = ""))
```

```{r include = FALSE}
nb_ab <- sum(grepl("_g", names(viparc))) - 1
```

The data frame contains `r nrow(viparc)` weeks of observation (rows) and
`r ncol(viparc)` variables (columns). The variables are

* **farm:** farm ID;
* **flock:** flock ID (in a given farm);
* **week:** week number (in a given flock of a given farm);
* **respiratory, ..., sudden_death:** presence (`TRUE`) / absence (`FALSE`) of 6 clinical signs in the flock;
* **amoxicillin_g, ..., unknown_g:** mass, in g, of the `r nb_ab` antimicriobial used in the flock;
* **completed:** boolean informing whether the flock is done (`TRUE`) or still ongoing (`FALSE`);
* **nb_chicken:** total number of chicken in the farm;
* **sampling:** boolean informing whether there is feces sampling during the week (`TRUE`) or not (`FALSE`).