---
title: "Testing the data from the CliRes database"
output:
  html_document:
    theme: cosmo
    toc: true
editor_options: 
  chunk_output_type: console
css: style.css
---

```{r general_options, include = FALSE}
knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  prompt = function(before, options, envir) {
    options(prompt = if (options$engine %in% c("sh", "bash")) "$ " else "> ")
  })

knitr::opts_chunk$set(margin = TRUE, prompt = TRUE, comment = "",
                      collapse = TRUE, cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 3.5,
                      fig.width = 4.24725, fig.retina = 2, fig.align = "center")

options(width = 250)
```

## Packages

Installing the required packages:

```{r}
required <- c("dplyr", "magrittr", "readr")
to_install <- setdiff(required, row.names(installed.packages()))
if (length(to_install)) install.packages(to_install)
```

Loading `magrittr`:

```{r}
library(magrittr)
```

## Loading the data

```{r}
viparc <- readr::read_csv("https://raw.githubusercontent.com/viparc/clires_data/master/data/viparc.csv",
                          col_types = paste(c("cii", rep("l", 6), rep("d", 45), "lil"), collapse = ""))
```

## Testing for missing weeks and missing values

No missing weeks:

```{r}
viparc %>%
    dplyr::group_by(farm, flock) %>%
    dplyr::arrange(week) %>%
    dplyr::summarise(a = length(unique(diff(week)))) %>%
    dplyr::ungroup() %>%
    dplyr::filter(a > 1)
```

Almost all the variables have missing values:

```{r}
sapply(viparc, function(x) any(is.na(x)))
```

But missing values concern only 2 weeks:

```{r}
weeks <- sort(unique(unlist(lapply(viparc, function(x) which(is.na(x))))))
viparc[weeks, c("farm", "flock", "week")]
```

## Checking for potential outliers

There does not seem te be any outliers in the number of chicken:

```{r}
hist(viparc$nb_chicken, col = "grey", main = NA, xlab = "number of chicken", ylab = "number of weeks")
```

There seems to be an error on the amount of oxytetracycline during one of the
weeks of observation:

```{r}
viparc %>%
  dplyr::select(dplyr::ends_with("_g")) %>% 
  unlist() %>%
  {.[. > 0]} %>%
  sort(TRUE) %>%
  head(20)
```

It concerns the following week:

```{r}
viparc %>%
  dplyr::filter(oxytetracycline_g > 200) %>%
  dplyr::select(farm, flock, week)
```

With this data point removed, it seems to make a little bit more sense:

```{r}
viparc %>%
  dplyr::select(dplyr::ends_with("_g")) %>% 
  unlist() %>%
  {.[. > 0]} %>%
  {.[. < 300]} %>%
  hist(0:200, col = "grey", main = NA, xlab = "antibiotic use (g)", ylab = "number of weeks")
```

